\vspace{-3em} \textbf{Occultation mapping of Io's surface in the near-infrared \\ I: Inferring static maps}

### 2021-06-13 paper/paper.tex
Improved sentence structure in the abstract

\begin{abstract}
Jupiter's moon Io is the most volcanically active body in the Solar System with hundreds of active volcanoes varying in intensity on different timescales.
Io has been observed during occultations by other Galilean moons and Jupiter since the 1980s, using high-cadence near infrared photometry. 
These observations encode a wealth of information about the volcanic features on its surface.
    We built a generative model for the observed occultations using the code \textsf{starry} which enables fast, analytic, and differentiable computation of occultation light curves in emitted and reflected light.
    Our probabilistic Bayesian model is able to recover known hotspots on the surface of Io using only two light curves and without any assumptions on the locations, shapes or the number of spots.
    The methods we have developed are also directly applicable to the problem of mapping the surfaces of stars and exoplanets. \href{https://github.com/fbartolic/volcano}{\color{linkcolor}\faGithub}
\end{abstract}

@@

To model the data covariance $\boldsymbol{\Sigma}$ we use use a Gaussian Process and we compute the likelihood using the fast Celerite method \citep{foreman-mackey2017} as implemented in the \textsf{celerite2} package \citep{foreman-mackey2017a,foreman-mackey2018}.
We use the simple (approximate) Mat\'ern 3/2 kernel function which is parametrized by two values, a standard deviation parameter $\sigma_\mathrm{GP}$ and a characteristic timescale parameter $\rho_\mathrm{GP}$.
The Mat\'ern 3/2 kernel is defined by
\begin{equation}
    k(\tau;\sigma_\mathrm{GP},
    \rho_\mathrm{GP})=\
    \sigma_\mathrm{GP}^{2}\left(1+\frac{\sqrt{3} \tau}{\rho_\mathrm{GP}}\right) \exp \left(-\frac{\sqrt{3} \tau}{\rho_\mathrm{GP}}\right)
    \quad,
\end{equation}
where $\tau=|t_n-t_m|$ and $\epsilon$ controls the quality of the approximation.
A single element of the data covariance matrix $\boldsymbol{\Sigma}$ is then 
\begin{equation}
    \boldsymbol{\Sigma}_{nm}=\sigma_n^2\delta_{nm} + k(\tau;\sigma_\mathrm{GP},\rho_\mathrm{GP})
    \quad,
    \label{eq:data_covariance_element}
\end{equation}
where $\sigma_n$ is the errorbar for the n-th data point.
Since we fit multiple independent light curves the total log likelihood is the sum of individual likelihoods defined in Equation~(\ref{eq:likelihood}).

@@

\appendix
\section{Horseshoe priors}
\label{app:horseshoe}
The Regularized Horseshoe prior \citep{piironen2017} is specifically designed for use in Bayesian sparse linear regression. 
It is a generalization of the Horseshoe prior introduced in \cite{carvalho2010a}.
The idea behind the Horseshoe prior is to set the scale for each regression coefficient (pixel) to a product of a global scale $\tau$ and a local scale $\lambda_i$ where $i$ indexes all the pixels.
The Horseshoe prior is defined hierarchically as 
\begin{equation}
\begin{aligned}
    &\tau  \sim \mathrm{Half}-\mathcal{C} \left(0, \tau_{0}\right)\\
    &\lambda_{i}  \sim \mathrm{Half}-\mathcal{C} (0,1) \\
    &p_{i}  \sim \mathcal{N}\left(0, \tau \lambda_{i}\right) 
\end{aligned}
    \quad,
    \label{eq:horseshoe}
\end{equation}
where $p_i$ are the pixels.
Each local scale parameter is drawn from a unit scale heavy tailed Half Cauchy distribution which allows for very large values of the pixels.
The global scale parameter is also a free parameter, drawn from a Half Cauchy distribution with the scale equal to $\tau_0$.
The Horseshoe prior is closely related to the spike-and-slab prior which is a mixture between a delta function prior at zero (\emph{spike}) and some other prior elsewhere (\emph{slab}).

The Regularized Horseshoe prior adds another level to Equation~(\ref{eq:horseshoe}) in order to allow fine tuned control of sparsity and to regularize very large values of coefficients in cases where the data is only weakly constraining.
\cite{piironen2017} show that the Regularized Horseshoe prior can be considered as a continuous counterpart of the spike-and-slab prior with a finite slab width $c$ whereas the Horseshoe prior resembles a spike-and-slab prior with a slab of infinite width.
The prior is defined by
\begin{equation}
\begin{aligned}
    &\tau  \sim \mathrm{Half}-\mathcal{C}\left(0, \tau_{0}\right)\\
    &c^{2}  \sim \mathrm{Inv}-\mathcal{G}\left(\frac{\nu}{2}, \frac{\nu}{2} s^{2}\right) \\
    &\overline{\lambda}_{i}  \sim \mathrm{Half}-\mathcal{C}(0,1)\\
    &\lambda_{i} =\frac{c \overline{\lambda}_{i}}{\sqrt{c^{2}+\tau^{2} \overline{\lambda}_{i}^{2}}} \\
    &p_{i}  \sim \mathcal{N}^+\left(0, \tau \lambda_{i}\right) 
\end{aligned}
    \label{eq:reg_horseshoe}
\end{equation}
Integrating out the slab scale $c$ implies a marginal $\mathrm{Student}-\mathit{t}\,(\nu,0,s)$ prior for pixels far from zero.
When pixels $p_i$ are close to zero ($\tau^2\overline{\lambda}_i^2\ll c^2$) we have $\lambda_i^2\rightarrow\overline{\lambda}_i^2$ and the prior approaches the original Horseshoe.
When pixels are far from zero ($\tau^2\overline{\lambda}_i^2\gg c^2$) then $\lambda_i^2\rightarrow c^2/\tau^2$ and the prior approaches $\mathcal{N}(0, c)$.

@@

When using the Regularized Horseshoe prior in a small data regime it is often necessary to use the non-centered parametrization to avoid funnels in the posterior which are often present in hierarchical models\footnote{\url{https://mc-stan.org/docs/2_26/stan-users-guide/reparameterization-section.html}}.
The purpose of this reparametrization is to reduce the dependence between the hyperparameters in the posterior.
To implement the non-centered parametrization we replace priors in Equation~\ref{eq:reg_horseshoe} with zero mean and unit variance priors and rescale them with deterministic transforms as follows
\begin{equation}
\begin{aligned}
    &\overline{\tau} \sim \mathrm{Half}-\mathcal{C}\left(0, 1\right)\\
    &\overline{c^{2}}  \sim \mathrm{Inv}-\mathcal{G}\left(\frac{\nu}{2}, 1\right) \\
    &\overline{\lambda}_{i}  \sim \mathrm{Half}-\mathcal{C}(0,1) \\
    &\overline{p}_{i}  \sim \mathcal{N}^+\left(0, 1\right)\\
    &\tau=\tau_0\overline{\tau}\\
    &c^2=\frac{\nu}{2}s^2\,\overline{c^2}\\
    &\lambda_{i} =\frac{c \overline{\lambda}_{i}}{\sqrt{c^{2}+\tau^{2} \overline{\lambda}_{i}^{2}}} \\
    &p_i = \tau\lambda_i\,\overline{p}_i
\end{aligned}
    \label{eq:reg_horseshoe_noncentered}
\end{equation}
We find that without using the non-centered parametrization the sampling is problematic and there are many divergences in the gradients of the parameters; with the non-centered parametrization there are no problems with sampling.

>>>>

\begin{abstract}
With hundreds of active volcanoes varying in intensity on different timescales, Jupiter's moon Io is the most volcanically active body in the Solar System.
Io has been observed from Earth using high-cadence near infrared photometry during occultations by Jupiter  and other Galilean moons since the 1980s.
These observations encode a wealth of information about the volcanic features on its surface.
    We built a generative model for the observed occultation light curves using the code \textsf{starry} which enables fast, analytic, and differentiable computation of occultation light curves in emitted and reflected light.
    Using this model we are able to recover surface emission maps of Io containing known volcanic hotspots, without having to make assumptions about the locations, shapes or the number of these hotspots.
    Our model is also directly applicable to the problem of mapping the surfaces of stars and exoplanets. \href{https://github.com/fbartolic/volcano}{\color{linkcolor}\faGithub}
\end{abstract}

@@

To model the data covariance $\boldsymbol{\Sigma}$ we use use a Gaussian Process and we compute the likelihood using the fast Celerite method \citep{foreman-mackey2017} as implemented in the \textsf{celerite2} package \citep{foreman-mackey2017a,foreman-mackey2018}.
We use the simple (approximate) Mat\'ern 3/2 kernel function which is parametrized by two values, a standard deviation parameter $\sigma_\mathrm{GP}$ and a characteristic timescale parameter $\rho_\mathrm{GP}$.
The Mat\'ern 3/2 kernel is defined by
\begin{equation}
    k(\tau;\sigma_\mathrm{GP},
    \rho_\mathrm{GP})=\
    \sigma_\mathrm{GP}^{2}\left(1+\frac{\sqrt{3} \tau}{\rho_\mathrm{GP}}\right) \exp \left(-\frac{\sqrt{3} \tau}{\rho_\mathrm{GP}}\right)
    \quad,
\end{equation}
where $\tau=|t_n-t_m|$.
A single element of the data covariance matrix $\boldsymbol{\Sigma}$ is then 
\begin{equation}
    \boldsymbol{\Sigma}_{nm}=\sigma_n^2\delta_{nm} + k(\tau;\sigma_\mathrm{GP},\rho_\mathrm{GP})
    \quad,
    \label{eq:data_covariance_element}
\end{equation}
where $\sigma_n$ is the errorbar for the n-th data point.
Since we fit multiple independent light curves the total log likelihood is the sum of individual likelihoods defined in Equation~(\ref{eq:likelihood}).

@@

\appendix
\section{Horseshoe priors}
\label{app:horseshoe}
The Regularized Horseshoe prior \citep{piironen2017} is specifically designed for use in Bayesian sparse linear regression. 
It is a generalization of the Horseshoe prior introduced in \cite{carvalho2010a}.
The idea behind the Horseshoe prior is to set the scale for each regression coefficient (pixel) to a product of a global scale $\tau$ and a local scale $\lambda_i$ where $i$ indexes all the pixels.
The Horseshoe prior is defined hierarchically as 
\begin{equation}
\begin{aligned}
    &\tau  \sim \mathrm{Half}-\mathcal{C} \left(0, \tau_{0}\right)\\
    &\lambda_{i}  \sim \mathrm{Half}-\mathcal{C} (0,1) \\
    &p_{i}  \sim \mathcal{N}\left(0, \tau \lambda_{i}\right) 
\end{aligned}
    \quad,
    \label{eq:horseshoe}
\end{equation}
where $p_i$ are the pixels and $\mathrm{Half}-\mathcal{C}$ is the Half Cauchy distribution.
Each local scale parameter is drawn from a unit scale heavy tailed Half Cauchy distribution which allows for very large values of the pixels.
The global scale parameter is also a free parameter, drawn from a Half Cauchy distribution with the scale equal to $\tau_0$.
The Horseshoe prior is closely related to the spike-and-slab prior which is a mixture between a delta function prior at zero (\emph{spike}) and some other prior elsewhere (\emph{slab}).

The Regularized Horseshoe prior adds another level to Equation~(\ref{eq:horseshoe}) in order to allow fine tuned control of sparsity and to regularize very large values of coefficients in cases where the data is only weakly constraining.
\cite{piironen2017} show that the Regularized Horseshoe prior can be considered as a continuous counterpart of the spike-and-slab prior with a finite slab width $c$ whereas the Horseshoe prior resembles a spike-and-slab prior with a slab of infinite width.
The prior is defined by
\begin{equation}
\begin{aligned}
    &\tau  \sim \mathrm{Half}-\mathcal{C}\left(0, \tau_{0}\right)\\
    &c^{2}  \sim \mathrm{Inv}-\mathcal{G}\left(\frac{\nu}{2}, \frac{\nu}{2} s^{2}\right) \\
    &\overline{\lambda}_{i}  \sim \mathrm{Half}-\mathcal{C}(0,1)\\
    &\lambda_{i} =\frac{c \overline{\lambda}_{i}}{\sqrt{c^{2}+\tau^{2} \overline{\lambda}_{i}^{2}}} \\
    &p_{i}  \sim \mathrm{Half}-\mathcal{N}\left(\tau \lambda_{i}\right) 
\end{aligned}
\quad,
    \label{eq:reg_horseshoe}
\end{equation}
where $\mathrm{Inv}-\mathcal{G}$ is the Inverse Gamma distribution and  $\mathrm{Half}-\mathcal{N}$ is the Half Normal distribution.
Integrating out the slab scale $c$ implies a marginal $\mathrm{Student}-\mathit{t}\,(\nu,0,s)$ prior for pixels far from zero.
When pixels $p_i$ are close to zero ($\tau^2\overline{\lambda}_i^2\ll c^2$) we have $\lambda_i^2\rightarrow\overline{\lambda}_i^2$ and the prior approaches the original Horseshoe.
When pixels are far from zero ($\tau^2\overline{\lambda}_i^2\gg c^2$) then $\lambda_i^2\rightarrow c^2/\tau^2$ and the prior approaches $\mathcal{N}(0, c)$.

@@

When using the Regularized Horseshoe prior in a small data regime it is often necessary to use the non-centered parametrization to avoid funnels in the posterior which are often present in hierarchical models\footnote{\url{https://mc-stan.org/docs/2_26/stan-users-guide/reparameterization-section.html}}.
The purpose of this reparametrization is to reduce the dependence between the hyperparameters in the posterior.
To implement the non-centered parametrization we replace priors in Equation~\ref{eq:reg_horseshoe} with zero mean and unit variance priors and rescale them with deterministic transforms as follows
\begin{equation}
\begin{aligned}
    &\overline{\tau} \sim \mathrm{Half}-\mathcal{C}\left(0, 1\right)\\
    &\overline{c^{2}}  \sim \mathrm{Inv}-\mathcal{G}\left(\frac{\nu}{2}, 1\right) \\
    &\overline{\lambda}_{i}  \sim \mathrm{Half}-\mathcal{C}(0,1) \\
    &\overline{p}_{i}  \sim \mathrm{Half}-\mathcal{N}\left(1\right)\\
    &\tau=\tau_0\overline{\tau}\\
    &c^2=\frac{\nu}{2}s^2\,\overline{c^2}\\
    &\lambda_{i} =\frac{c \overline{\lambda}_{i}}{\sqrt{c^{2}+\tau^{2} \overline{\lambda}_{i}^{2}}} \\
    &p_i = \tau\lambda_i\,\overline{p}_i
\end{aligned}
    \label{eq:reg_horseshoe_noncentered}
\end{equation}
We find that without using the non-centered parametrization the sampling is problematic and there are many divergences in the gradients of the parameters; with the non-centered parametrization there are no problems with sampling.
